---
title: "Human Activity Recognition : Practical Machine Learning (Coursera)"
author: "Debjyoti Paul"
date: "Sunday, December 21, 2014"
output: html_document
keep_md: yes
bibliography: bibliography.bib
---

# Abstract
In this project, a predictive system has been modelled using Random Forest to infer on the qualitative aspect of Human Activity Recognition using weight lifting exersise dataset [@qualitative]. It will help the users to have a real time feedback on the quality of excetution of the exersise based on sensor data. 

# Data
This project is based on the Weight Lifting Exercises(WLE) Dataset from Human Activity Recognition project of GroupWare <http://groupware.les.inf.puc-rio.br/har>.

# Initilization

The seed is set to enable reproduction. Libraries required are loaded

```{r}
set.seed(111)
library(caret)
library(randomForest)

```

# Getting Data

The training and testing data are downloaded from the url provided.

```{r}
tr.file <- 'pml-training.csv'
te.file <- 'pml-test.csv'
tr.url  <- 'http://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv'
te.url  <- 'http://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv'

download.file(tr.url, tr.file, method='auto')
download.file(te.url, te.file, method='auto')
```

# Cleaning Data
Data is read from .csv format. All the coloumns with inappropriate values or index information are removed.

```{r}
trdata = read.csv(file=tr.file, header=TRUE, sep=",")
tedata = read.csv(file=te.file, header=TRUE, sep=",")
trdata = subset( trdata, select = -1 )
tedata = subset( tedata, select = -1 )
nums <- sapply(trdata, is.numeric)
cleandata <- trdata[,nums]
training <- cleandata[ , colSums(is.na(cleandata)) == 0]
```

# Partition Data

The cleaned data is partitioned into Training and Testing dataset.

```{r}
trainingIndex  <- createDataPartition(trdata$classe, p=.75, list=FALSE)
trainData <- training[ trainingIndex,]
testData  <- training[-trainingIndex,]
```
# Training
The predctive model is trained using the training data.

## Method
The **Random Forests** method [@random], which applies bagging to tree learners. The **Bootstap Aggregating** (**bagging**) method is described in [@bagging]


## Train
The model is trained using **train** function. The "**method = "rf"**" specifies the random forest method.
```{r}
modFit <- train(trdata[trainingIndex,]$classe ~ .,method = "rf", data = trainData)
```


# Result

The predictive model after training and the Training result
```{r}
modFit

print(modFit$finalModel)
```

```{r, echo=FALSE}
plot(modFit)
```
An instance of a tree in the random forest.
```{r}
getTree( modFit$finalModel , k = 3)
```

Confusion matrix and statistics depicting the Performance of the model on the testing data(testData)

```{r}
pred <- predict( modFit, newdata = testData )
confusionMatrix(pred,trdata[-trainingIndex,]$classe)
```

```{r}
plot( varImp(modFit) )
```


#References